---
title: "part 2 chi square lab!"
author: "Conner Jainese"
date: "11/28/2021"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Red in Data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

# Get counts of lizards by species

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), EAST (east tarbush site)
- Keep observations where tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))

```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)

```

Return counts in contingency table format for chi squared test

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
  

```

Is there a significant association between site and tail breakage?

# Convert to a contingency table 

Use `column_to_rowname()`

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square
Are tail breakage and site independent?
Do we have enough evidence to reject the null hypothesis of independence?
Is there a significant effect of site on tail breakage?
- basically all the same question

```{r}
lizard_x2 <- chisq.test(lizard_ct)

lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```
- Using a significance level of p = 0.05, we can reject the null hypothesis that site and tail breakage are independent.
- There is a significant association between site and tail breakage. 
- Site and lizard tail breakage are not independent

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).






